TacticToe: Learning the Game of Tactics

### 2018-01-26 tactictoe_jv.tex

\subsection{The problem}\todo{Maybe should }

Definitions of common concepts: goals, tactics, theorems, terms , etc

\subsection{Contributions}
This project is an extension of the work described in \cite{tgckju-lpar17}. We 
explain here the additional contributions and their effect across our system:

\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. So, our recording is more complete and fully automatized.
%\item Existing tactic orthogonalization on recorded data is made stronger by 
%also including the internal ATP \metis in the orthogonalization 
%competition. Because we manage a database of theorems for argument 
%prediction, we apply here a pruning step to remove easy theorems.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently of the tactic. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.

@@

\section{Proof Search}\label{sec:proofsearch}
\todo
{
TG:
describe the prior policy, the evaluation, 
the node selection algorithm,
the node extension (+ case of solved goal) + backpropagation.
add description of end of the algorithm.
}
Despite the best efforts of the prediction algorithms, the selected tactic may 
not solve the current goal, proceed in the wrong direction
or even loop. For that reason, the prediction needs to be accompanied by a  
proof search mechanism that allows for backtracking and 
can choose which proof tree to extend next and in which direction.
Our search algorithm is a Monte Carlo tree search (MCTS) algorithm that relies 
on prior evaluation and policy functions. Those priors are learned through 
direct supervised learning in the case of the policy and via the data 
accumulated during orthogonalization for the evaluation. This MCTS do not rely 
on roll outs. The 
policy and evaluation are estimated by the simple supervised k-nn algorithm 
here instead of neural networks trained through reinforcement learning in 
AlphaGo Zero~\cite{}.

@@

%
%\subsection{Advantage of Monte Carlo Tree Search over A*-search}
%- Take into account all estimation along the proof and not only the last leaf.
%- Easy to link conjunction estimation   together.
-

\section{ATP Integration}\todo{TG + JU + CK: rework and review}
General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined 
power of tactics, we would like to combine the \tactictoe search with 
stronger general purpose provers such as the one in \holyhammer for 
\holfour~\cite{tgck-cpp15}.

@@

Rather than to rely on the underlying programming language, we parse the
original theory file containing proof scripts into an
abstract syntax tree (AST).\todo{reason: reading the proof as if you were a 
human ==> better for the machine learning algorithm (learn the right thing)}

@@

\subsection{Globalization}

@@

%\subsubsection {Unfolding locally declared values}
To transfer \sml values as code across theories in a conservative manner, we 
need to 
globalize \sml identifiers. To this end, we replace each identifier by some 
\sml code containing all the information to obtain the value of this identifier 
without knowing the context. The result is a standalone \sml code executable in 
any \holfour theories which will be interpreted in the same way in the absence 
of side effects. Two techniques are helpful to produce such code. The first one 
is to prefix global identifiers by the module (and sub-modules) in which they 
were declared. And the second one is to replace local identifiers 
by their definitions. Prefixing and 
unfolding are applied recursively in 
the definitions.

@@

\subsection{Storing Declaration}

@@

\subsection{Globalizing Expression}

@@

\subsection{Recording tactic calls}

@@

\subsection{Storing Training Data}
We describe here the format preferred to store and load the 
data recorded by the function \texttt{save}. 
During saving, we add fields to the input pair $(tactic,goal)$ to record 
precomputed information in order to make the prediction faster. The final 
feature vector is the tuple $(tactic,time,goal,goal\_list,features)$.
The time $time$ taken by $tactic$ is not currently used by our 
predictor. The list of goals $goal list$ is the output of $tactic$ on 
$goal$. And the features $features$ of the goal $goal$ are extracted and hashed 
as described in Section~\ref{sec:features}.

>>>>

\subsection{The problem}\todo{ }

How a user works with an ITP and how this gives rise to a machine learning problem.

Two ways of representing a formulas.

Definition of a tactic in \holfour. In this research, we will take the definition
that a tactic takes a goal and returns a list of goals.

\subsection{Contributions}
This work is an extension of the work described in \cite{tgckju-lpar17}.
There we proposed the idea of ... and achieved ...
The additional contributions and their effect on our system are:

\todoi{add prediction of 4 kinds of things}
\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. So, our recording is more complete and fully automatized.
%\item Existing tactic orthogonalization on recorded data is made stronger by 
%also including the internal ATP \metis in the orthogonalization 
%competition. Because we manage a database of theorems for argument 
%prediction, we apply here a pruning step to remove easy theorems.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently of the tactic. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.

@@

\section{Proof Search}\label{sec:proofsearch}

Despite the best efforts of the prediction algorithms, the selected tactic may 
not solve the current goal, proceed in the wrong direction
or even loop. For that reason, the prediction needs to be accompanied by a  
proof search mechanism that allows for backtracking and 
can choose which proof tree to extend next and in which direction.
Our search algorithm is a Monte Carlo tree search (MCTS) algorithm that relies 
on prior evaluation and policy functions. Those priors are learned through 
direct supervised learning in the case of the policy and via the data 
accumulated during orthogonalization for the evaluation. This MCTS do not rely 
on roll outs. The 
policy and evaluation are estimated by the simple supervised k-nn algorithm 
here instead of neural networks trained through reinforcement learning in 
AlphaGo Zero~\cite{}.

@@

%
%\subsection{Advantage of Monte Carlo Tree Search over A*-search}
%- Take into account all estimation along the proof and not only the last leaf.
%- Easy to link conjunction estimation   together.

\subsection{ATP Integration}
General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined 
power of tactics, we would like to combine the \tactictoe search with 
stronger general purpose provers such as the one in \holyhammer for 
\holfour~\cite{tgck-cpp15}.

@@

Rather than relying on the underlying programming language, we parse the
original theory file containing proof scripts into an
abstract syntax tree (AST).\todo{reason: reading the proof as if you were a 
human ==> better for the machine learning algorithm (learn the right thing)}

@@

@@

\subsection{Globalization}
%\subsubsection {Unfolding locally declared values}
To transfer \sml values as code across theories in a conservative manner, we 
need to 
globalize \sml identifiers. To this end, we replace each identifier by some 
\sml code containing all the information to obtain the value of this identifier 
without knowing the context. The result is a standalone \sml code executable in 
any \holfour theories which will be interpreted in the same way in the absence 
of side effects. Two techniques are helpful to produce such code. The first one 
is to prefix global identifiers by the module (and sub-modules) in which they 
were declared. And the second one is to replace local identifiers 
by their definitions. Prefixing and 
unfolding are applied recursively in 
the definitions.

@@

\subsubsection{Declaration}

@@

\subsubsection{Expression}

@@

\subsection{Tactic wrap}

@@

\subsection{Data Transfer}
We describe here the format preferred to save and load the 
data recorded by the function \texttt{save}. 
During saving, we add fields to the input pair $(tactic,goal)$ to record 
precomputed information in order to make the prediction faster. The final 
feature vector is the tuple $(tactic,time,goal,goal\_list,features)$.
The time $time$ taken by $tactic$ is not currently used by our 
predictor. The list of goals $goal list$ is the output of $tactic$ on 
$goal$. And the features $features$ of the goal $goal$ are extracted and hashed 
as described in Section~\ref{sec:features}.
